Plots

### 2025-06-18 src/plots.Rmd
Cleaner plots

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = NULL,
        y_title = NULL,
        text_mid_size = 2 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 3 * text_scale,
        text_end_size = 4 * text_scale,
        text_hjust = 0.1,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 2.5,
        min_points_size = 2,
        max_points_size,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -2)),
        ...
    ) {

@@

    # Add metric value to model name
    df_plot <- df_plot %>%
        mutate( 
            model_label = paste0(model_name, " (", round(metric, 2), ")")
        )

@@

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_point( # points
            shape = 21,
            color = "black",
            aes(fill = .data[[color]]),
            size = mid_points_size
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, " (", round(min_metric, 2), ")")
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_start_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(model_max, " (", round(max_metric, 2), ")")
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Plot perf of models using trend lines
plot_perf_trend <- function(
        df_long,
        f = calc_pccc,
        group = "age_range",
        text_scale = 1,
        title = NULL,
        x_title = NULL,
        y_title = NULL,
        title_size = 10,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        legend_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_direction = "horizontal",
        legend_spacing = unit(0, "pt")
    ) {

@@

    # Plot trend lines for each model
    out <- ggplot(
        df_plot,
        aes(
            x = .data[[group]],
            y = metric,
            color = model_name,
            group = model_name
        )
    ) +
        geom_line(linewidth = 0.75, alpha = 0.4) + 
        geom_line(
            data = df_plot %>%
                group_by(.data[[group]]) %>%
                summarize(
                    metric = mean(metric),
                    model_name = "Mean Across Models"
                ),
            aes(color = model_name),
            se = F,
            linewidth = 1.25
        ) +
        scale_color_manual(values = default_colors) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.title = element_blank(),
            legend.text = element_text(size = legend_text_size),
            legend.position = legend_position, 
            legend.direction = legend_direction,
            legend.box.spacing = legend_spacing,
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)

@@

# Model full name
gpt3_name <- "GPT-3.5"
gpt4_name <- "GPT-4"
isva_name <- "InSilicoVA"
iva5_name <- "InterVA-5"

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = factor(model_name, levels = model_name)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "CSMF Accuracy") %>%
            factor(., levels = rev(c("PCCC", "CSMF Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", "")
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 3,
    sort_group = F,
    x_limits = c(0.2, 0.8),
    legend_margin = margin(-8, -2, -2, -2),
    x_expand = waiver()
)

# Save plot
ggsave(
    "../manuscript/fig-pccc.pdf",
    fig_pccc,
    width = 170,
    height = 60,
    units = "mm",
    dpi = 300
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult, 12+ Years (n=", n_adult, ")"),
    x_title =  "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = -1,
    text_scale = 0.8
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child, 28 Days to 11 Years (n=", n_child, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_scale = 0.95
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 3,
    text_end_size = 2.25,
    text_scale = 0.95
)

>>>>

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = NULL,
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 3 * text_scale,
        text_end_size = 4 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 2.5,
        min_points_size = 2,
        max_points_size = 3,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -2)),
        ...
    ) {

@@

    # Add metric value to model name
    df_plot <- df_plot %>%
        mutate( 
            model_label = paste0(model_name, ": ", round(metric, 2))
        )

@@

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_point( # points
            shape = 21,
            color = "black",
            aes(fill = .data[[color]]),
            size = mid_points_size
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, ": ", round(min_metric, 2))
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_start_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(model_max, ": ", round(max_metric, 2))
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_fill_discrete(
            labels = x_color_labels
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Plot perf of models using trend lines
plot_perf_trend <- function(
        df_long,
        f = calc_pccc,
        group = "age_range",
        text_scale = 1,
        title = NULL,
        x_title = NULL,
        y_title = NULL,
        title_size = 10,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        legend_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_direction = "horizontal",
        legend_spacing = unit(0, "pt"),
        x_color_labels = c(
            "G3" = "GPT-3.5",
            "G4" = "GPT-4",
            "ISV" = "InSilicoVA",
            "I5" = "InterVA-5"
        )
    ) {

@@

    # Plot trend lines for each model
    out <- ggplot(
        df_plot,
        aes(
            x = .data[[group]],
            y = metric,
            color = model_name,
            group = model_name
        )
    ) +
        geom_line(linewidth = 0.75, alpha = 0.4) + 
        geom_line(
            data = df_plot %>%
                group_by(.data[[group]]) %>%
                summarize(
                    metric = mean(metric),
                    model_name = "Mean Across Models"
                ),
            aes(color = model_name),
            se = F,
            linewidth = 1.25
        ) +
        scale_color_manual(values = default_colors, labels = x_color_labels) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.title = element_blank(),
            legend.text = element_text(size = legend_text_size),
            legend.position = legend_position, 
            legend.direction = legend_direction,
            legend.box.spacing = legend_spacing,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)

@@

# Model full name
gpt3_name <- "G3"
gpt4_name <- "G4"
isva_name <- "ISV"
iva5_name <- "I5"

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = model_name %>%
            case_when(
                . == "G3" ~ "GPT-3.5",
                . == "G4" ~ "GPT-4",
                . == "ISV" ~ "InSilicoVA",
                . == "I5" ~ "InterVA-5",
                .default = .
            ) %>%
            factor(., levels = .)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "CSMF Accuracy") %>%
            factor(., levels = rev(c("PCCC", "CSMF Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", "")
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 2,
    sort_group = F,
    x_limits = c(0.2, 0.8),
    legend_margin = margin(-8, -2, -2, -2),
    x_expand = waiver(),
    text_scale = 1.25
)

# Save plot
ggsave(
    "../manuscript/fig-pccc.pdf",
    fig_pccc,
    width = 170,
    height = 75,
    units = "mm",
    dpi = 300
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult, 12+ Years (n=", n_adult, ")"),
    x_title =  "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = -1,
    text_scale = 1
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child, 28 Days to 11 Years (n=", n_child, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 4,
    text_scale = 1.15
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 3,
    text_scale = 1.25
)
